---
title: "Workshop 7 R GeoDa"
author: "Anastasia Bernat"
date: "11/19/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tips
Use Cmd + option + i to embed a R code (```{r}) - R Markdown cheat sheet: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf

## Load and Explore Data 

```{r}
library(rgeoda)
library(geodaData)
library(sf)

guerry_sf <- geodaData::guerry # classic social science foundational study by Andre-Michel Guerry on crime, suicide, etc.
summary(guerry_sf)
names(guerry_sf)
str(guerry_sf)
dim(guerry_sf)
head(guerry_sf)
st_crs(guerry_sf)

```

## Plot Data

```{r}
plot(guerry_sf)
```

## Metadata

"Description

Classic social science foundational study by Andre-Michel Guerry on crime, suicide, literacy and other “moral statistics” in 1830s France. Data from the R package Guerry (Michael Friendly and Stephane Dray)."

```{r}
?guerry
```

## Exploring Data More

Can read shapefiles, csv's, etc. files directly with geoda_open
If want to work on attributes work in sf and then move here after wards

sf_to_geoda(guerry_sf, with_table = TRUE) # optimizing for autocorrelation, throw away all attributes except for geometry, but if you add with_table- = TRUE then you get all your attirbutes back. 

```{r}
guerry <- sf_to_geoda(guerry_sf, with_table = TRUE) # geoda object
class(guerry) # rgeoda object, not normal data frame 

# s-4 method - use a $ and tab and functions pop out

guerry$table
guerry$field_names
guerry$map_type
guerry$n_cols # can do this? 


```

## Calculate Queen Weights

```{r}
queen_w <- queen_weights(guerry)
queen_w
queen_w$density # series of methods or functions with the $
queen_w$is_symmetric
queen_w$SaveToFile("queen_weights.gal", "guerry", "CODE_DE", guerry$GetIntegerCol("CODE_DE")) 
```
 gal (geometrics algorithm lab - simple text file that continas, for each obs, the num of neighbors and their identifiers)
 vs. gwt # path and, name of ID column
 
```{r}
class(queen_w) # weight object 

rook_w <- rook_weights(guerry)
rook_w
# knn_weights(guerry)

?queen_weights
```

## Local Moran

```{r}
crm_prp <- guerry_sf$Crm_prp
crm_prp1 <- guerry$GetIntegerCol("crm_prp")
lisa <- local_moran(queen_w, crm_prp) # apply local Moran statistics on attribute column 

class(lisa) # LISA object
lms <- lisa$GetLISAValues()
lisa$GetPValues()

# Go through clustering next week

lisa_clusters <- lisa$GetClusterIndicators()
lisa_colors <- lisa$GetColors()
lisa_labels <- lisa$GetLabels()
lisa_colors
plot(st_geometry(guerry_sf), 
     col=sapply(lisa_clusters, function(x){return(lisa_colors[[x+1]])}), 
     border = "#333333", lwd=0.2)
title(main = "Local Moran Map of Crm_prp")
legend('bottomleft', legend = lisa_labels, fill = lisa_colors, border = "#eeeeee")

# local_geary(queen_w, crm_prp) # ??? Used to calculate the statistics -  chapter 5
# geary_crmprp<- lisa$GetClusterIndicators()
```




